---
title: "PSTAT 126"
subtitle: "Lab 3"
author: "Roupen Khanjian"
date: "Spring 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(palmerpenguins) # Palmer Archipelago (Antarctica) Penguin Data
```

\tableofcontents

Dataset: Adelie and Gentoo Penguins, with the same question as last section:  

- **Can we predict body mass in grams by a penguins bill length in mm?**

```{r}
data("penguins")

penguins_noChinstrap <- penguins %>% 
  filter(species != "Chinstrap") %>% 
  drop_na(bill_length_mm, body_mass_g)

model <- lm(body_mass_g ~ bill_length_mm , data = penguins_noChinstrap)
summary(model)

```


## Simple Linear Regression Model Assumptions

1) The relationship between each $Y_n$ and each $x_n$, respectively, is linear.  **L**inearity 
2) Errors have **E**qual variance. Var($Y_n$) =  $\sigma^2$ for every $n$ (homoscedasticity)
3) Errors are **N**ormally distributed 
4) Errors are **I**ndependent 

* Can use the acronym **L.I.N.E.** to help you remember. 

#### Graphically checking the normality assumption 

&nbsp;
&nbsp;

##### QQ - plot

&nbsp;
&nbsp;
```{r}
par(mfrow = c(1, 2))

plot(model, which = 2) # QQ

e <- residuals(model) # Residuals

qqnorm(e) # QQ
qqline(e)

```
&nbsp;
&nbsp;

\newpage

##### Histogram of residuals

&nbsp;
&nbsp;

```{r}
par(mfrow = c(1, 1))
resid_model <- tibble(residuals = residuals(model))

ggplot(data = resid_model,
       aes(x = residuals)) +
  geom_histogram(color = "white",
            fill = "seagreen",
            bins = 15) +
  labs(x = "Residuals",
       title = "Histogram of Residuals") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

```

\newpage

## Confidence interval for new observations

* Here we use $x_0 = 50$ bill length (mm)

**95% Confidence Interval for new observation**

```{r}
n <- nrow(penguins_noChinstrap) # number of observations
x <- penguins_noChinstrap$bill_length_mm # predictor variable
y <- penguins_noChinstrap$body_mass_g # response variable
x_bar <- mean(x) # mean of bill_length_mm
y_bar <- mean(y) # mean of body_mass_g
Sxx <- sum((x - x_bar) ^ 2)
sigma_hat <- summary(model)$sigma # Residual Standard Error (RSE)
Yhat_50 <- # predicated body mass when bill length is 50 mm
  as.numeric(coef(model)[1] + coef(model)[2] * 50) 
y_hat <- fitted(model) # fitted values

spe_50 <- sigma_hat*sqrt(1 + 1/n + (50 - x_bar)^2/Sxx) # se of y_hat(x_0)
t_pct <- qt(p = 0.975, df = n - 2) # t-statistic
CI_95 <- c(Yhat_50 - spe_50*t_pct, Yhat_50 + spe_50*t_pct) 
CI_95

predict(model, newdata = data.frame(bill_length_mm = 50), 
        level = 0.95, interval = 'prediction')

```

### Visualizing confidence interval bands

```{r}
ngrid <-  274
grid <-  seq(min(x), max(x), length = ngrid)
new <- data.frame(bill_length_mm = grid)
p2 <- predict(model, new, se.fit = TRUE, interval = "prediction", level = 0.95)

# tibble for plot
conf_pred_tib <- tibble(x,y, y_hat, new = new$bill_length_mm, 
                        UL_p = p2$fit[,3], LL_p = p2$fit[,2])
```

\newpage

```{r}
# Plot
ggplot(data = conf_pred_tib) +
  geom_point(aes(x = x, y = y), color = "darkgreen", alpha = 0.8, size = 2) + # data points
  geom_line(aes(x = x, y = y_hat), color = "blue") + # Fitted line
  geom_line(aes(x = new, y = UL_p), color = "red", linetype = "dashed") + # upper bound
  geom_line(aes(x = new, y = LL_p), color = "red", linetype = "dashed")+ # lower bound
  scale_x_continuous(breaks = seq(30, 60, by = 10)) +
  scale_y_continuous(breaks = seq(2000, 7000, by = 1000)) +
  labs(x = "bill length (mm)",
       y = "body mass (grams)",
       title = "95% Prediction bands") +
  theme_minimal() +
  theme()


```

\newpage

```{r}

ggplot(data = conf_pred_tib) +
  geom_point(aes(x = x, y = y), color = "darkgreen", alpha = 0.8, size = 2) + # data points
  geom_line(aes(x = x, y = y_hat), color = "blue") + # Fitted line
  geom_line(aes(x = new, y = UL_p), color = "red", linetype = "dashed") + # upper bound
  geom_line(aes(x = new, y = LL_p), color = "red", linetype = "dashed")+ # lower bound
  geom_vline(aes(xintercept = 50), size = 1, linetype = "dotted") + # x0 value
  scale_x_continuous(breaks = seq(30, 60, by = 10)) +
  scale_y_continuous(breaks = seq(2000, 7000, by = 1000)) +
  labs(x = "bill length (mm)",
       y = "body mass (grams)",
       title = "95% Confidence and Prediction bands with x0 value") +
  theme_minimal() +
  theme()

predict(model, newdata = data.frame(bill_length_mm = 50), 
        level = 0.95, interval = 'predict')

```

\newpage

## Coefficient of Determination $R^2$

* A goodness-of-fit measure

$$R^2 = 1 - \frac{RSS}{S_{yy}}$$
```{r}
b0 <- summary(model)$coef[1,1] # Intercept
b1 <- summary(model)$coef[2,1] # Slope
y_hat <- b0 + b1*x # Fitted values
e <- y - y_hat # Residuals

Syy <- sum((y - y_bar)^2)

r_2 <- 1 - (sum(e^2)/Syy)
r_2

summary(model)$r.squared

r <- cor(x,y)
r^2
```

Notes on $R^2$

* Always between 0 and 1
* Can interpret as $R^2 \times 100$ percent of the variation in Y is explained by the variation in the predictor x. 



