---
title: "PSTAT 126"
subtitle: "Lab 10"
author: "Roupen Khanjian"
date: "Spring 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```



```{r}
library(faraway) # Functions and Datasets for Books by Julian Faraway 
library(alr4) # Data to Accompany Applied Linear Regression 4th Edition
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(patchwork) # The Composer of Plots
library(GGally) # Extension to 'ggplot2'
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data
library(palmerpenguins) # Palmer Archipelago (Antarctica) Penguin Data
library(broom) # Convert Statistical Objects into Tidy Tibbles
library(lmtest) # Testing Linear Regression Models
library(ballr) # Access to Current and Historical Basketball Data
```

\tableofcontents
\newpage

## Review of Linear Regression Assumptions

#### Linear Regression Model Assumptions
&nbsp;

1) The relationship between each $Y_n$ and each $x_n$, respectively, is linear.  **L**inearity 
2) Errors have **E**qual variance. Var($Y_n$) =  $\sigma^2$ for every $n$ (homoscedasticity)
3) Errors are **N**ormally distributed 
4) Errors are **I**ndependent 

* Can use the acronym **L.I.N.E.** to help you remember. 

### Model 1: Not Satisfying Assumptions

```{r}
nba_data1 <- NBAPerGameStatistics(season = 2021)

nba_data <- distinct(nba_data1, player, .keep_all = TRUE)  %>% 
  filter(pts > 2) %>% 
  filter(g > 10) %>% 
  filter(pos %in% c("PG", "SG")) %>% 
  mutate(pos = factor(pos))
```

&nbsp;
```{r}
ggplot(data = nba_data,
       aes(x = pts)) +
  geom_histogram(fill = "dodgerblue", color = "white",
                 bins = 15) +
  geom_rug(color = "firebrick", alpha = 0.7) +
  labs(title = "Histogram of Response Variable") +
  theme_bw()
```
&nbsp;

```{r}
model_nba1 <- lm(pts ~ pos + tov + trb, data = nba_data)
summary(model_nba1)

par(mfrow = c(1,2))
plot(model_nba1, which = 1, add.smooth = F) # resid vs fit
plot(model_nba1, which = 2) # qqplot
shapiro.test(resid(model_nba1))  # resid vs fit
bptest(model_nba1) # test for homoscedasticity
dwtest(model_nba1) # test for autocorrelation
gqtest(model_nba1) # test for homoscedasticity
```

\newpage

### Model 2: Satisfying Assumptions

```{r}
nba_data <- distinct(nba_data1, player, .keep_all = TRUE)  %>% 
  filter(pts > 8) %>% 
  filter(g > 20) %>% 
  filter(pos %in% c("PF", "C")) %>% 
  mutate(pos = factor(pos))
```

&nbsp;
```{r}
par(mfrow = c(1,1))
ggplot(data = nba_data,
       aes(x = mp)) +
  geom_histogram(fill = "dodgerblue", color = "white",
                 bins = 15) +
  geom_rug(color = "firebrick", alpha = 0.7) +
  labs(title = "Histogram of Response Variable") +
  theme_bw()
```

&nbsp;

```{r}
model_nba2 <- lm(mp ~ stl + tov, data = nba_data)
summary(model_nba2)

par(mfrow = c(1,2))
plot(model_nba2, which = 1, add.smooth = F) # resid vs fit
plot(model_nba2, which = 2) # qqplot
shapiro.test(resid(model_nba2)) # resid vs fit
bptest(model_nba2) # test for homoscedasticity
dwtest(model_nba2) # test for autocorrelation
gqtest(model_nba2) # test for homoscedasticity
```

\newpage

## Generalized Linear Models: Logistic Regression

* Last week we used species from the penguins dataset as a categorical predictor variable. Today let's use it as our response variable in a logistic regression example. 

#### Data
&nbsp;

```{r}

penguins_noNA_no_Gentoo <- penguins %>% 
  drop_na() %>% 
  filter(species != "Gentoo" )%>% 
  mutate(species = fct_drop(species))

slice_sample(penguins_noNA_no_Gentoo, n = 5)

levels(penguins_noNA_no_Gentoo$species)

```

#### EDA
&nbsp;

```{r}
mycolors <- c("salmon1", "seagreen3")

par(mfrow = c(1,1))
ggplot(data = penguins_noNA_no_Gentoo, 
       aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = mycolors) +
  facet_wrap(~ species) +  
  labs(x = "Body Mass (g)",
       y = "Flipper length (mm)",
       color = "Sex") +
  theme_bw()
```


#### Logistic Regression Model
&nbsp;

```{r}
model_species <- glm(species ~ body_mass_g + flipper_length_mm + sex, 
                     data = penguins_noNA_no_Gentoo, 
                     family = "binomial")

summary(model_species)

tidy(model_species)

coef(model_species)

```

#### Likelihood Ratio Test
&nbsp;

```{r}

model_species_smaller <- glm(species ~ flipper_length_mm + body_mass_g, 
                     data = penguins_noNA_no_Gentoo, 
                     family = "binomial")

model_species_larger <- glm(species ~ body_mass_g + flipper_length_mm + sex + bill_depth_mm, 
                     data = penguins_noNA_no_Gentoo, 
                     family = "binomial")

anova(model_species_smaller,
      model_species_larger,
      test = "LRT")
```

\newpage

#### Prediction
&nbsp;

What's the probabilty that a male penguin weighing 4150 grams with a flipper length of 205 mm is a Chinstrap?

&nbsp;

```{r}
new_data <- data.frame(
  body_mass_g = 4150,
  flipper_length_mm = 205,
  sex = "male")

predict(model_species, newdata = new_data, se.fit = TRUE,
        type = "response")$fit
```

What's the probability that a female penguin weighing 4950 grams with a flipper length of 190 mm is a Chinstrap?

&nbsp;

```{r}
new_data2 <- data.frame(
  body_mass_g = 4950,
  flipper_length_mm = 190,
  sex = "female")

predict(model_species, newdata = new_data2, se.fit = TRUE,
        type = "response")$fit

```

```{r}

model_fitted <- augment(model_species, type.predict = "response")

slice_sample(model_fitted, n = 5)[1:5]

ggplot(data = model_fitted, aes(x = flipper_length_mm, y = .fitted)) +
  geom_point(aes(color = species)) +
  labs(x = "Flipper length (mm)",
       y = "Probability of Chinstrap") +
  theme_minimal()

```


