---
title: "PSTAT 126"
subtitle: "Lab 7 Part 2"
author: "Roupen Khanjian"
date: "Spring 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
library(faraway) # Functions and Datasets for Books by Julian Faraway 
library(alr4) # Data to Accompany Applied Linear Regression 4th Edition
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(lmtest) # Testing Linear Regression Models
```

\tableofcontents


* Note on Homework. Make sure you are knitting as you go along. Don't wait until you do all your analysis and then knit. 

\newpage

## ANOVA using the `aov` function

```{r}
data("chickwts")
glimpse(chickwts)
table(chickwts$feed)

ggplot(data = chickwts, 
       aes(x = feed, y = weight)) +
  geom_boxplot(aes(fill = feed),
               show.legend = FALSE) +
  theme_minimal() +
  theme(axis.text = element_text(face = "bold",
                                 size = 10))
```

\newpage

```{r}
chick_wt_aov <- aov(weight ~ feed, data = chickwts)
chick_wt_aov
summary(chick_wt_aov)

anova(lm(weight ~ feed, data = chickwts))

```

\newpage

## Data example

```{r}
data("starwars")
head(starwars)

starwars_subset <- starwars %>% # subset data
  select(height, mass) %>% # select mass as response and height as predictor
  drop_na() %>% # remove missing values
  filter(mass < 500) # filter for mass under 500 kg

ggplot(data = starwars_subset,
       aes(x = height, y = mass)) +
  geom_point(color = "forestgreen",
             alpha = 0.85, size = 2) +
  theme_minimal()

model_sw <- lm(mass ~ height, starwars_subset)
summary(model_sw)

ggplot(data = starwars_subset,
       aes(x = height, y = mass)) +
  geom_point(color = "forestgreen",
             alpha = 0.85, size = 2) +
  geom_abline(aes(intercept = coef(model_sw)[1],
                  slope = coef(model_sw)[2]),
              color = "firebrick", size = 1.25) +
  theme_minimal()
```

* check model assumptions graphically and using formal tests. 

```{r}
plot(model_sw, which = 1, add.smooth = F)
plot(model_sw, which = 2)

bptest(model_sw)
shapiro.test(resid(model_sw))
shapiro.test(residuals(model_sw))
```

* assumptions are violated 


* try response transformation

```{r}
bc <- boxCox(model_sw)
bc$x[which.max(bc$y)]
```

lets try to log transform response

```{r}
model_sw2 <- lm(log(mass) ~ height, data = starwars_subset)
summary(model_sw2)
ggplot(data = starwars_subset,
       aes(x = height, y = log(mass))) +
  geom_point(color = "forestgreen",
             alpha = 0.85, size = 2) +
  geom_abline(aes(intercept = coef(model_sw2)[1],
                  slope = coef(model_sw2)[2]),
              color = "firebrick", size = 1.25) +
  theme_minimal()
```

```{r}
plot(model_sw2, which = 1, add.smooth = F)
plot(model_sw2, which = 2)

bptest(model_sw2)
shapiro.test(resid(model_sw2))
```


* Should we add a polynomial term maybe? Let's see!

```{r}
model_sw3 <- lm(log(mass) ~ height + I(height^2), data = starwars_subset)
summary(model_sw3)
```

```{r}
anova(model_sw2, model_sw3)
```

* should we transform the predictor variable?

```{r}
invTranPlot(log(mass) ~ height, data = starwars_subset,
            lambda = c(0, 1), optimal = TRUE)
```

* From the plot above can try a log transformation of predictor variable. 

```{r}
model_sw4 <- lm(log(mass) ~ log(height), data = starwars_subset)
summary(model_sw4)

```

```{r}
plot(model_sw4, which = 1, add.smooth = F)
plot(model_sw4, which = 2)

bptest(model_sw4)
shapiro.test(resid(model_sw4))
```

* Normality assumption is violated. 

**Therefore, the most appropriate model for this data would be model 2** 

```{r}
summary(model_sw2)
```








