---
title: "PSTAT 126"
subtitle: "Lab 8"
author: "Roupen Khanjian"
date: "Spring 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```



```{r}
library(faraway) # Functions and Datasets for Books by Julian Faraway 
library(alr4) # Data to Accompany Applied Linear Regression 4th Edition
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(leaps) # Regression Subset Selection
library(patchwork) # The Composer of Plots
```


## Model selection

#### Data from Faraway book (Chapter 10)

* Suppose the intercept is included in the model. For the remaining p - 1 covariates (predictors) , they could be in the model or out. Then in total we have $2^{p-1}$ choices. When p = 8, we have 128 potential models (not counting interaction or polynomial terms!).  


```{r}
data(state)
statedata <- data.frame(state.x77, row.names = state.abb)
head(statedata)
lmod <- lm(Life.Exp ~ ., statedata)
```

\newpage
```{r}
summary(lmod)
```



```{r}
b <- regsubsets(formula(lmod),
                data=statedata)
rs <- summary(b) 
rs$which # for each model of size p+1, chooses the model with the lowest RSS value.
```

```{r}
# plot(rs$rss ~ I(1:7), ylab="RSS", 
#      xlab="Number of Predictors", main = "RSS vs # of Predictors" )
```

\newpage
```{r}
r1 <- ggplot(data = data.frame(rs$rss), aes(x = 2:8, y = rs$rss)) +
  geom_point(colour = "red", size = 1.5) +
  geom_label(aes(label= round(rs$rss, 3)), size = 3, nudge_y = 2 ) +
  scale_x_continuous(breaks = seq(2,8,1)) +
  ylim(22, 37) +
  labs(x = "Number of parameters", y = "RSS",
       title = "RSS") +
  theme_minimal()

r2 <- ggplot(data = data.frame(rs$rsq), aes(x = 2:8, y = rs$rsq)) +
  geom_point(colour = "red", size = 1.5) +
  geom_label(aes(label = round(rs$rsq, 4)), size = 3, nudge_y = 0.02) +
  scale_x_continuous(breaks = seq(2, 8, 1)) +
  ylim(0.6, 0.77) +
  labs(x = "Number of parameters", y = "R^2",
       title = "R^2") +
  theme_minimal()

r1 / r2

```

\newpage

**Now we introduce information criteria for model selection.**

* **Akaike’s Information Criterion (AIC)**

$$AIC = n log(RSS) - n log(n) + 2p  \\
    = n log(RSS/n) + 2p$$
    
* In AIC $k = 2$

* **Bayesian Information Criterion (BIC)**

$$BIC = n log(RSS) - n log(n) + p(log(n))  \\
    = n log(RSS/n) + p(log(n))$$  
    
* In BIC $k =  log(n)$

#### Notes on AIC/BIC

* BIC penalizes larger models more heavily and so will tend to prefer smaller models in comparison to AIC.

* The goal is to identify a subset of predictors such that AIC or BIC are minimized.

```{r}
n <- nrow(statedata)
AIC <- n*log(rs$rss/n) + (2:8)*2
BIC <- n*log(rs$rss/n) + (2:8)*(log(n))
# plot(BIC~ I(1:7), ylab="BIC", xlab="Number of Predictors")
# plot(AIC ~ I(1:7), ylab="AIC", xlab="Number of Predictors")
```

\newpage
```{r}
a1 <- ggplot(data = data.frame(AIC), aes(x = 2:8, y = AIC)) +
  geom_point(colour = "red", size = 1.5) +
  geom_label(aes(label= round(AIC, 2)), size = 3, nudge_y = 2 ) +
  scale_x_continuous(breaks = seq(2,8,1)) +
  ylim(-30, -11) +
  labs(x = "Number of parameters", y = "AIC") +
  ggtitle("AIC") +
  theme_minimal()

b1 <- ggplot(data = data.frame(BIC), aes(x = 2:8, y = BIC)) +
  geom_point(colour = "red", size = 1.5) +
  geom_label(aes(label= round(BIC, 2)), size = 3, nudge_y = 2) +
  scale_x_continuous(breaks = seq(2,8,1)) +
  ylim(-20, -4) +
  labs(x = "Number of parameters", y = "BIC") +
  ggtitle("BIC") +
  theme_minimal()

a1 / b1

```

\newpage
#### Model Selection

* Forward selection
  + Start with no variables (just intercept)
  + Add one variable at a time according to some criterion
  + Stop when no more variables should be added
* Backward selection
  + Start with a Full model with all possible predictors
  + Remove one variable at a time according to some criterion
  + Stop when no more variables should be dropped

##### Forward selection using p-values
* Let $\alpha = 0.10$ be our stopping criteria. 

```{r}
mod0 <- lm(Life.Exp ~ 1, statedata)
add1(mod0, ~.+Population+Income+Illiteracy+Murder+HS.Grad+Frost+Area, test = "F")

mod1 <- update(mod0, ~.+Murder)
add1(mod1, ~.+Population+Income+Illiteracy+HS.Grad+Frost+Area, test = "F")

mod2 <- update(mod1, ~.+HS.Grad)
add1(mod2, ~.+Population+Income+Illiteracy+Frost+Area, test = "F")

mod3 <- update(mod2, ~.+Frost)
add1(mod3, ~.+Population+Income+Illiteracy+Area, test = "F")

mod4 <- update(mod3, ~.+Population)
add1(mod4, ~.+Income+Illiteracy+Area, test = "F")

summary(mod4)

```

\newpage
`step` function allows you to choose a model using AIC as the information criteria. Can use forward or backward selection. 

##### Forward selection using AIC values
```{r}
mod0 <- lm(Life.Exp ~ 1, statedata)
mod.upper <- lm(Life.Exp ~ ., statedata)
step(mod0, 
     scope = list(lower = mod0, upper = mod.upper),
     direction = "forward") # Forward (start with intercept)
```

##### Backward selection using AIC values

```{r}
lmod <- lm(Life.Exp ~ ., statedata)
step(lmod, direction = "backward") # backward is the default direction in R

```




