---
title: "PSTAT 126"
subtitle: "Lab 5"
author: "Roupen Khanjian"
date: "Spring 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
library(faraway) # Functions and Datasets for Books by Julian Faraway 
library(alr4) # Data to Accompany Applied Linear Regression 4th Edition
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(GGally) # Extension to 'ggplot2'
library(palmerpenguins) # Palmer Archipelago (Antarctica) Penguin Data
library(Lahman) # Sean 'Lahman' Baseball Database 
library(patchwork) # The Composer of Plots
```

\tableofcontents

## Linear Regression Model Assumptions

1) The relationship between each $Y_n$ and each $x_n$, respectively, is linear.  **L**inearity 
2) Errors have **E**qual variance. Var($Y_n$) =  $\sigma^2$ for every $n$ (homoscedasticity)
3) Errors are **N**ormally distributed 
4) Errors are **I**ndependent 

* Can use the acronym **L.I.N.E.** to help you remember. 

#### How to test these assumptions?

* Linearity and Constant Variance = Residuals vs. Fitted plot
* Normality = QQ plot

##### Examples

**Baseball example**

```{r}
df1 <- Batting %>% 
  filter(yearID == "2017" & # stats from 2017
           lgID == "NL" & # only from the NL
           AB > 100 & 
           AB < 600 ) # Only At Bats between 100-600.  

dim(df1)

Runs <- df1$R
At_Bats <- df1$AB

model_Runs <- lm(Runs ~ At_Bats)
summary(model_Runs)

ggplot() +
  geom_point(aes(x = At_Bats, y = Runs), color = "blue", 
             alpha = 0.85, size = 2.25) +
  geom_abline(aes(intercept = coef(model_Runs)[1],
                  slope = coef(model_Runs)[2]), 
                  color = "red") +
  labs(x = "At Bats",
       y = "Runs",
       title = "Plot with fitted values") +
  theme_minimal()

```



```{r}
par(mfrow = c(1, 2))

e1 <- resid(model_Runs) # Residuals
y_hat1 <- fitted(model_Runs) # Fitted Values

plot(model_Runs, which = 1, add.smooth = F) # Resid vs. Fit

plot(y_hat1, e1,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted plot")  # Resid vs. Fit
abline(0,0)

plot(model_Runs, which = 2) # QQ

qqnorm(e1) # QQ
qqline(e1)

```

From the QQ-plot we can see that the residuals have a heavy-tailed distribution. 


**Penguins Example**

```{r}

penguins_noChinstrap <- penguins %>% 
  filter(species != "Chinstrap") %>% 
  drop_na(bill_length_mm, body_mass_g)

model <- lm(body_mass_g ~ bill_length_mm , data = penguins_noChinstrap)
summary(model)

par(mfrow = c(1, 2))

e1 <- resid(model) # Residuals
y_hat1 <- fitted(model) # Fitted Values

plot(model, which = 1,add.smooth = F) # Resid vs. Fit

plot(y_hat1, e1,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted plot")  # Resid vs. Fit
abline(0,0)

plot(model, which = 2) # QQ

e <- residuals(model) # Residuals

qqnorm(e) # QQ
qqline(e)

par(mfrow = c(1, 1))
resid_model <- tibble(residuals = residuals(model))

ggplot(data = resid_model,
       aes(x = residuals)) +
  geom_histogram(color = "white",
            fill = "seagreen",
            bins = 15) +
  labs(x = "Residuals",
       title = "Histogram of Residuals") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

```


**Soccer Example**
```{r}

head(worldcup)

ggpairs(worldcup[,c("Time", "Passes", "Position")], lower.panel = NULL)

model_3 <- lm(Time ~ Passes + Position, worldcup)

plot(model_3, which = 1, add.smooth = F)
plot(model_3, which = 2, add.smooth = F)

```

\newpage

## F-Tests

#### Global F-test


&nbsp;
&nbsp;

```{r}
worldcup1 <- worldcup %>%
  select(-c(Saves, Team)) %>% 
  filter(Passes > 1 ) %>% 
  filter(Time > 90) %>% 
  filter(Position != "Goalkeeper") %>% 
  filter(Tackles > 0) %>% 
  filter(Shots > 0)
```

&nbsp;
&nbsp;

```{r}
g1 <- ggplot(data = worldcup1,
       aes(x = Passes)) +
  geom_histogram(color = "white",
            fill = "seagreen",
            bins = 10) +
  labs(x = "Passess",
       title = "Histogram of Passess") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

g2 <- ggplot(data = worldcup1,
       aes(x = log(Passes))) +
  geom_histogram(color = "white",
            fill = "seagreen",
            bins = 10) +
  labs(x = "Log Passess",
       title = "Histogram of log Passes") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

g1 / g2


model_Full <- lm(log(Passes) ~ Time  + Shots + Tackles, data = worldcup1)

# Global F test
model_null <- lm(log(Passes) ~ 1, data = worldcup1) 
anova(model_null, model_Full)
summary(model_Full)
```

Lets define our variables the following way:

$log(\hat{Y}) = log(\text{Passes})$  
$x_1 = Time$  
$x_2 = Shots$  
$x_3 = Tackles$  

$$log(\hat{Y}) = 3.4004649  + 0.0040118x_1 - 0.0149556x_2 + 0.0208857x_3$$
$$\therefore \text{If we want to get a value for }\hat{Y}$$
$$\hat{Y} = e^{3.4004649  + 0.0040118x_1 - 0.0149556x_2 + 0.0094249x_3}$$

For example if a player has played 100 minutes, has attempted 3 shots, and has made 10 tackles, then we would expect this player to make about
`r  exp(3.4004649  + 0.0040118*100 - 0.0149556*3 + 0.0094249*10)` passes.

```{r}
exp(3.4004649  + 0.0040118*100 - 0.0149556*3 + 0.0094249*10)
```

\newpage
#### Partial F-tests

* Testing a Subset of Slope Parameters Equal 0

R = Reduced model
F = Full model
N = number of observations
M = number of predictor variables

$$F^{*}= \frac{\frac{RSS(R) - RSS(F)}{(N - M - 1)_R - (N - M - 1)_F}}{\frac{RSS(F)}{(N - M - 1)_F}}$$


Model without `Shots` or `Tackles`

```{r}


model_reduced <- lm(log(Passes) ~ Time, data = worldcup1) 
anova(model_reduced, model_Full)

summary(model_Full)
```

\newpage

Model without `Shots` 

```{r}
model_reduced2 <- lm(log(Passes) ~ Time + Tackles, data = worldcup1) 
anova(model_reduced2, model_Full)

summary(model_Full)

```

\newpage

## Adjusted R^2

```{r}

statedata <- data.frame(state.x77, row.names = state.abb)
head(statedata)
lmod <- lm(Life.Exp ~ ., statedata)
summary(lmod)

library(leaps) # Regression Subset Selection
b <- regsubsets(formula(lmod),
                data=statedata)
rs <- summary(b) 
rs$which # for each model of size p+1, chooses the model with the lowest RSS value.

p1 <- ggplot(data = data.frame(rs$rsq), aes(x = 2:8, y =rs$rsq)) +
  geom_point(colour = "red", size = 1.5) +
  geom_label(aes(label= round(rs$rsq, 4)), size = 3, nudge_y = 0.01 ) +
  scale_x_continuous(breaks = seq(2,8,1)) +
  labs(x = "Number of parameters", y = "R^2",
       title = "R^2") +
  theme_minimal()

p2 <- ggplot(data = data.frame(rs$adjr2), aes(x = 2:8, y =rs$rsq)) +
  geom_point(colour = "red", size = 1.5) +
  geom_label(aes(label= round(rs$adjr2, 4)), size = 3, nudge_y = 0.01 ) +
  scale_x_continuous(breaks = seq(2,8,1)) +
  labs(x = "Number of parameters", y = "Adjusted R^2",
       title = "Adjusted R^2") +
  theme_minimal()

p1 / p2

```



