---
title: "PSTAT 126"
subtitle: "Lab 2"
author: "Roupen Khanjian"
date: "Spring 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(palmerpenguins) # Palmer Archipelago (Antarctica) Penguin Data
```

\tableofcontents

## Computing OLS estimators in simple linear regression (without lm())

#### Dataset: Adelie and Gentoo Penguins   

- **Question: Can we predict body mass in grams by a penguins bill length in mm?**

```{r}
data("penguins")

penguins_noChinstrap <- penguins %>% 
  filter(species != "Chinstrap") %>% 
  drop_na(bill_length_mm, body_mass_g)

str(penguins_noChinstrap)
summary(penguins_noChinstrap)

# plot of data
ggplot(data = penguins_noChinstrap,
       aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point(color = "dodgerblue", alpha = 0.9, size = 1.75) +
  theme_minimal()


```

```{r}
x <- penguins_noChinstrap$bill_length_mm
y <- penguins_noChinstrap$body_mass_g
```

\newpage

First obtain means of $x$ and $y$
```{r}
x_bar <- mean(x) 
y_bar <- mean(y)
```

$$S_{xx} :\Sigma_{i = 1}^n (x_i - \bar{x})^2$$ 
```{r}
Sxx <- sum((x - x_bar)^2)
Sxx
```
$$S_{yy} :\Sigma_{i = 1}^n (y_i - \bar{y})^2$$ 
```{r}
Syy <- sum((y - y_bar)^2)
Syy
```

$$S_{xy} :\Sigma_{i = 1}^n (x_i - \bar{x})(y_i - \bar{y})$$
```{r}
Sxy <- sum((x - x_bar)*(y - y_bar))
Sxy
```

$$\hat\beta_1 = S_{xy}/S_{xx}$$
```{r}
b1 <- Sxy / Sxx
b1
```


$$\hat\beta_0 = \bar{y} - \hat\beta_1\bar{x}$$
```{r}
b0 <- y_bar - b1*x_bar
b0
```

$$\hat{Y} = \hat\beta_0 + \hat\beta_1x$$
```{r}
y_hat <- b0 + b1*x
```

**Estimation of Residuals**  
$$e_i = y_i - \hat{y}$$  
  
```{r}
e <- y - y_hat
```

$$\hat\sigma^2 = \frac{1}{N-2}\Sigma_{n = 1}^n e_n^2$$


```{r}
n <- length(y)
sigma_2_hat <- sum(e^2) / (n-2)
sigma_2_hat
sqrt(sigma_2_hat) # Residual Standard Error (RSE)
```


\newpage
## The lm() function

```{r}
model <- lm(body_mass_g ~ bill_length_mm , data = penguins_noChinstrap)

```


```{r}
summary(model)
```

\newpage

```{r}
coef(model) # Estimates for b0 and b1
model$coefficients
```


```{r}
head(residuals(model)) # residuals
head(fitted(model)) # y_hat values

summary(residuals(model)) # First line in summary output.
```


```{r}
# Standard errors
summary(model)$coef[,2]
coef(summary(model))[, "Std. Error"]
```


```{r}
# p-values for intercept and slope
summary(model)$coef[,4]
```

p-values for t-test and F-test in SLR are identical.

```{r}
summary(model)$sigma^2
```

\newpage

## Confidence Intervals for intercept and slope estimates 

Can calculate a 90% confidence interval by entering values into formula:

* **Intercept**

$$\hat\beta_0 \pm (t_{\alpha/2, N-2}\boldsymbol{SE}(\hat\beta_0))$$

* **Slope**

$$\hat\beta_1 \pm (t_{\alpha/2, N-2}\boldsymbol{SE}(\hat\beta_1))$$

```{r}
n <- length(x)
sigma_2_hat <- sum(e^2) / (n-2)
sigma_hat <- sqrt(sigma_2_hat) 
Sxx <- sum((x - x_bar)^2)

se_b0 <- sqrt(sigma_2_hat*(1/n + 
                             (x_bar^2)/Sxx)) # se of intercept
se_b1 <-  sqrt(sigma_2_hat/Sxx) # se of slope

t_pct <- qt(p = 0.95, df = n - 2) # t-statistic
```


```{r}
CI_b0_90 <-  c(b0 - t_pct*se_b0, b0 + t_pct*se_b0) # 90% CI for b0
CI_b1_90 <-  c(b1 - t_pct*se_b1, b1 + t_pct*se_b1) # 90% CI for b1
CI_b0_90
CI_b1_90
```


Can also use the `confint` function
```{r}
?confint
confint(model, level = 0.95) # 95% CI

confint(model, level = 0.90) # 90% CI

```

\newpage

## Hypothesis Testing

**Hypothesis testing of $\hat\beta_0, \hat\beta_1$**

Want to test:  
  
$H_0:\hat\beta_0 = 0$ vs. $H_1:\hat\beta_0 \neq 0$  
$H_0:\hat\beta_1 = 0$ vs. $H_1:\hat\beta_1 \neq 0$  
Let $\alpha = 0.05$  
```{r}
t_b0 <- (b0-0)/se_b0
t_b1 <- (b1-0)/se_b1
t_b0
t_b1

```

* For distributions in R, p stands for "probability", the cumulative distribution function (c. d. f.).  
```{r}

p0 <- 2*(1 - pt(abs(t_b0), df = n-2))
p1 <- 2*(1 - pt(abs(t_b1), df = n-2))

alpha <- 0.05
p0 > alpha
p1 > alpha

p0
p1

```
Reject Null Hypothesis for both $\hat\beta_0, \hat\beta_1$


\newpage

## Plots


```{r}
plot(body_mass_g ~ bill_length_mm , data = penguins_noChinstrap,
     main = "Plot with fitted values")
abline(model, col = "Red")
```


```{r}
ggplot(data = penguins_noChinstrap) +
  geom_point(aes(x = bill_length_mm, y = body_mass_g), color = "dodgerblue", alpha = 0.95) +
  geom_abline(aes(intercept = model$coefficients[[1]],
                  slope = model$coefficients[[2]]), 
                  color = "red") +
  labs(x = "bill length (mm)",
       y = "body mass (grams)",
       title = "Plot with fitted values") +
  theme_minimal()

```






